---
title: 'Final Project: Summary Statistics'
author: "David Aarhus"
date: "4/23/2020"
output: word_document
---

This should include a summary table of means, max, mins, and standard deviations; data transformations performed for feature engineering, as well as at least five plots revealing interesting patterns to be studied. These must be motivated by the analysis – they cannot just be random plots.

Needs improvement:

Clarify for the final report:


What are the sensitivity and specificity in the classification model?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) #removing all variables
```

#Libraries
```{r}
library("tidyverse")
library("ggplot2")
library("caret")
library("doBy")
```

```{r}
rm(list = ls()) #removing all variables
```

#Loads in Data
```{r}
suicide <- read.csv("/Users/DavidAarhus/Documents/310 R/Datasets/suicide.csv")
names(suicide)
```

#Data Transformation
```{r}
#remove meaning less columns (HDI.for.year)
suicide$HDI.for.year <- NULL
suicide$country.year <- NULL
#create new column (gdpM, gdpB, suicide_hvy)
suicide$gdpM <- suicide$gdp_for_year..../1e+6
suicide$gdpB <- suicide$gdp_for_year..../1e+9
suicide$pop100000 <- suicide$population*1e-5
suicide$suicide_rate_per100000capita <- suicide$pop100000/suicide$suicides_no

#identified whether a country was suicide heavy or not by first running summary statistics on the dataset to find the average suicide count from each country. 
suicide$suicide_hvy <- ifelse(suicide$suicide_rate_per100000capita > 0.56, 1, 0)


NaRV.omit <- function(x){
  ## Author: Rene Locher
  ## Version: 2005-10-17

  if (is.vector(x)) {
    if (is.numeric(x)) x <- na.omit(x[is.finite(x)]) else
    x <- na.omit(x)
  } else if (is.factor(x)) {
    x <- na.omit(x)
  } else if (is.data.frame(x)) {
    x.num <- as.matrix(x[,sapply(x, is.numeric)])
    ri <- (!apply(x.num, MARGIN = 1,
                  function(x) sum(is.infinite(x))>0) &
           !apply(x, MARGIN = 1, function(x) sum(is.na(x))>0))
    x <- x[ri,,drop=FALSE]
    ## class omit is incompatible with class data.frame
    ## attributes(x) <- c(attributes(x),list(na.action=which(!ri),class="omit"))
  } else if (is.matrix(x)) {
    if(is.numeric(x)) {
      ri <- !apply(x, MARGIN = 1,
                   function(x) {sum(is.na(x)|is.infinite(x))>0})
      x <- x[ri,,drop=FALSE]
      attributes(x) <- c(attributes(x),list(na.action=which(!ri),class="omit"))
    } else x <- na.omit(x)
  } else {
    warning("'x' is neither a vector, nor a factor, nor a data.frame nor a matrix. \n'x' is returned unchanged\n")
    return(x)
  }
  return(x)
}## NaRV.omit

suicide <- NaRV.omit(suicide)

```


#Summary tables
```{r}
summary(suicide)

#sum(suicide$suicide_hvy==1)
#4368
```


```{r}
summaryBy(suicides_no ~ country, suicide, FUN = mean)
```
```{r}
summaryBy(gdpB ~ country, suicide, FUN = mean)
```

```{r}
summaryBy(suicides_no ~ sex, suicide, FUN = mean)
```

#GGPLOTS
```{r}
ggplot(suicide, aes(generation)) +
  geom_bar(aes(fill = sex)) +
  labs(x = " ", y = "Count", title = "Suicide Totals vs Generations")
```

```{r}
ggplot(suicide, aes(suicides.100k.pop, gdpB)) +
  geom_point(aes(color = age)) +
  labs(x = "Suicides per 100k pop ", y = "GDP (Billions)", title = "GDP (Billions) vs Suicides per 100k pop")

```

```{r}
ggplot(suicide, aes(year, suicides_no)) +
  geom_point(aes(color = sex)) +
  labs(x = "Year", y = "Suicide Count", title = "Suicides vs Year")

```

```{r}
ggplot(suicide, aes(population, suicides.100k.pop )) +
  geom_point(aes(color = generation)) +
    labs(x = "Population", y = "Suicides per 100k pop", title = "Suicides per 100k pop Vs Population")
```

```{r}
ggplot(suicide, aes(year, suicides.100k.pop)) +
  geom_point(aes(color = generation)) +
  labs(x = "Year", y = "Suicides per 100k pop", title = "Suicides per 100k pop Vs Year")

```