---
title: "Final Project_Models"
author: "David Aarhus"
date: "4/29/2020"
output: word_document
---

Students must upload to Canvas a compiled Rmarkdown document that shows one model(s) estimated against the data. Code and replication files are required.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r}
library("tidyverse")
library("ggplot2")
library("caret")
library("doBy")
library("glmnet")
library("glmnetUtils")
library("leaps")
library("plotROC")

```

```{r}
rm(list = ls()) #removing all variables
```

#Loads in Data
```{r}
suicide <- read.csv("/Users/DavidAarhus/Documents/310 R/Datasets/suicide.csv")
```

#Data Transformation
```{r}
#remove meaning less columns (HDI.for.year)
suicide$HDI.for.year <- NULL
suicide$country.year <- NULL
#create new column (gdpM, gdpB, suicide_hvy)
suicide$gdpM <- suicide$gdp_for_year..../1e+6
suicide$gdpB <- suicide$gdp_for_year..../1e+9
suicide$pop100000 <- suicide$population*1e-5
suicide$suicide_rate_per100000capita <- suicide$pop100000/suicide$suicides_no

#identified whether a country was suicide heavy or not by first running summary statistics on the dataset to find the average suicide rate for each country. This then allowed me to identify the average suicide rate for the dataset. After finding this I used a suicide rate just a tad higher than the average to identify whether a country was suicide heavy (in relation to the worldwide average)
suicide$suicide_hvy <- ifelse(suicide$suicide_rate_per100000capita > 0.56, 1, 0)

NaRV.omit <- function(x){
  ## Author: Rene Locher
  ## Version: 2005-10-17

  if (is.vector(x)) {
    if (is.numeric(x)) x <- na.omit(x[is.finite(x)]) else
    x <- na.omit(x)
  } else if (is.factor(x)) {
    x <- na.omit(x)
  } else if (is.data.frame(x)) {
    x.num <- as.matrix(x[,sapply(x, is.numeric)])
    ri <- (!apply(x.num, MARGIN = 1,
                  function(x) sum(is.infinite(x))>0) &
           !apply(x, MARGIN = 1, function(x) sum(is.na(x))>0))
    x <- x[ri,,drop=FALSE]
    ## class omit is incompatible with class data.frame
    ## attributes(x) <- c(attributes(x),list(na.action=which(!ri),class="omit"))
  } else if (is.matrix(x)) {
    if(is.numeric(x)) {
      ri <- !apply(x, MARGIN = 1,
                   function(x) {sum(is.na(x)|is.infinite(x))>0})
      x <- x[ri,,drop=FALSE]
      attributes(x) <- c(attributes(x),list(na.action=which(!ri),class="omit"))
    } else x <- na.omit(x)
  } else {
    warning("'x' is neither a vector, nor a factor, nor a data.frame nor a matrix. \n'x' is returned unchanged\n")
    return(x)
  }
  return(x)
}## NaRV.omit

suicide <- NaRV.omit(suicide)
```

```{r}
set.seed(310)
train_indx <- sample(1:nrow(suicide), 0.70 * nrow(suicide), replace=FALSE)
suicide_train <- suicide[train_indx, ]
suicide_test <- suicide[-train_indx, ]
```

#Linear Regression
```{r}
# We are using a linear model to see the affect of certain variables on the number of suicides in different countries
Linear_mod <- lm(suicide_rate_per100000capita ~ ., suicide_train)
summary(Linear_mod)
```

```{r}
#This linear model helps point out the factors other than countries that play a key role in the amount of suicides
Linear_mod_noCountries <- lm(suicide_rate_per100000capita ~ year + sex 
                             + age + generation + country
                             + gdpB + gdpM, suicide_train)
summary(Linear_mod_noCountries)
```

# Linear Regression Predictions
```{r}
preds_train1 <- predict(Linear_mod_noCountries, newdata = suicide_train)
preds_test1 <- predict(Linear_mod_noCountries, newdata = suicide_test)



preds_train1_df <- data.frame(true = suicide_train$suicide_rate_per100000capita,
                              pred = predict(Linear_mod_noCountries, newdata = suicide_train),
                              resid = suicide_train$suicide_rate_per100000capita - predict(Linear_mod_noCountries, newdata =  suicide_train))



preds_test1_df <- data.frame(true = suicide_test$suicide_rate_per100000capita,
                              pred = predict(Linear_mod_noCountries, newdata = suicide_test),
                              resid = suicide_test$suicide_rate_per100000capita - predict(Linear_mod_noCountries, newdata = suicide_test))
```

## Plot predictions
```{r}
ggplot(preds_train1_df, aes(x=pred, y=resid)) +
  geom_point() + 
  geom_smooth(se=FALSE)
```

```{r}
ggplot(preds_test1_df, aes(x=pred, y=resid)) +
  geom_point() + 
  geom_smooth(se=FALSE)
```

## MSE for train and test
```{r}
RMSE(preds_train1_df$pred, preds_train1_df$true)
```

```{r}
RMSE(preds_test1_df$pred, preds_test1_df$true)
```

##Logistic Regression
```{r}
# this logistic regression model helps identify what variables contribute to a country being "suicide heavy" (365 or more suicides per year)
logit_fit <- glm(suicide_hvy ~ year + sex + age + generation + country, family = binomial, suicide_train)
summary(logit_fit)
```

##Logistic Regression Predictions
```{r}
preds_train <- data.frame(scores = predict(logit_fit, type = "response"), suicide_train)
preds_train <- data.frame(class_preds05 = ifelse(preds_train$scores > 0.5, 1, 0), preds_train)

preds_test <- data.frame(scores = predict(logit_fit, newdata = suicide_test, type = "response"), suicide_test)
preds_test <- data.frame(class_preds05 = ifelse(preds_test$scores > 0.5, 1, 0), preds_test)

```

##Train confusion matrix
```{r}
table(preds_train$suicide_hvy,preds_train$class_preds05)
```

##Test confusion matrix
```{r}
table(preds_test$suicide_hvy, preds_test$class_preds05)
```

#Train ROC Plot
```{r}
train_ROC <- ggplot(preds_train, aes(m = scores, d = suicide_hvy)) +
  geom_roc(cutoffs.at = c(0.99,0.90,0.70,0.3,0.1,0.01))
train_ROC
```

##Test ROC Plot
```{r}
test_ROC <- ggplot(preds_test, aes(m = scores, d = suicide_hvy)) +
  geom_roc(cutoffs.at = c(0.99,0.90,0.70,0.3,0.1,0.01))
test_ROC
```

##Train AUC score
```{r}
calc_auc(train_ROC)
```

#Test AUC score
```{r}
calc_auc(test_ROC)
```

#Forward Stepwise Model
```{r}
# This forward stepwise model helps identify which variables contribute the most to the amount of suicides. 
forward_mod <- regsubsets(suicide_rate_per100000capita ~ year + sex 
                 + gdp_for_year.... + gdp_per_capita....
                 + generation + gdpM + gdpB, data = suicide_train, nvmax = 6, method = "forward")
summary(forward_mod)
```
